Single Site, Single Disturbance

### 2020-10-03 SingleSite_SingleDisturb.Rmd
Iterative JAGS filter (single site, single disturb) working!

```{r}
FilterSSSD <- "
model{
  N0 ~ dnorm(mu0,pa0)  ## initial conditions
  ## Forecast
  muN <- N0 + R*N0*(1-N0/K)
  muD <- 1
  D ~ dbern(p)
  mu <- D[t]*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.LFP = matrix(NA,3,nt)
Xl = data$mu0
pl = data$pa0
aqp = bqp = arp = brp = 0.1
mu.m = 0
p.m  = 1e-6
for(t in 1:nt){
  ## Forecast + Analysis
  update = list(Y=Y[t],mu0=Xl,pa0=pl,aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],mu.m=mu.m[t],tau.m=p.m[t])
  mod <- jags.model(file=textConnection(LinFilterParam),
                   data=update,
                  n.adapt=1000,n.chains=3,
                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("m","r","q","X"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"X"])
  pl  = 1/var(dat[,"X"])
  CI.LFP[,t] = quantile(dat[,"X"],c(0.025,0.5,0.975))
  mq = mean(dat[,"q"])
  vq = var(dat[,"q"])
  aqp[t+1] = mq*mq/vq
  bqp[t+1] = mq/vq
  mr = mean(dat[,"r"])
  vr = var(dat[,"r"])
  arp[t+1] = mr*mr/vr
  brp[t+1] = mr/vr
  mu.m[t+1] = mean(dat[,"m"])
  p.m[t+1] = 1/var(dat[,"m"])
}

@@

## Todo

-   Figure on what happens when a SDA model that doesn't include disturbance encounters disturbance
-   More detail on what a model w/ disturbance but EnKF does during disturbance
-   Get the iterative filter working to see how it compares to the other two
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA

>>>>

```{r}
FilterSSSD <- "
model{
  N ~ dnorm(muIC,paIC)  ## initial conditions
  ## Forecast
  muN <- N + R*N*(1-N/K)
  muD ~ dnorm(mu0,pa0)
  D ~ dbern(p)
  mu <- D*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Xl = nIC         ## set IC
pl = 1/n.sd^2
Dbar = 0
#mu0 = data$mu0   ## set disturbed state
#pa0 = data$pa0
#aqp = data$aq
#bqp = data$bq
#arp = data$ar
#brp = data$br
for(t in 1:NT){
  ## Forecast + Analysis
  update = list(Y=data$Y[t],
                muIC=Xl,paIC=pl,
                mu0=data$mu0,pa0=data$pa0,
                #aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],
                R=r, K=K,p = p.dist, q = 1/n.sd^2, r=1/obs.sd^2     ## hard coding these to start
                )
  mod <- jags.model(file=textConnection(FilterSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)#,
#                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"n"])
  pl  = 1/var(dat[,"n"])
  CI.fN[,t] = quantile(dat[,"n"],c(0.025,0.25,0.5,0.75,0.975))
  CI.fD[,t] = quantile(dat[,"D"],c(0.025,0.25,0.5,0.75,0.975))
  Dbar[t] = mean(dat[,"D"])
#  mu0[t+1] = mean(dat[,"muD"])
#  pa0[t+1] = 1/var(dat[,"muD"])

#  mq = mean(dat[,"q"])
#  vq = var(dat[,"q"])
#  aqp[t+1] = mq*mq/vq
#  bqp[t+1] = mq/vq
#  mr = mean(dat[,"r"])
#  vr = var(dat[,"r"])
#  arp[t+1] = mr*mr/vr
#  brp[t+1] = mr/vr
}

@@

```{r}
plot(0,0,xlim=xlim,ylim=c(0,1))
ecoforecastR::ciEnvelope(time,CI.fD[1,],CI.fD[5,],col=col.alpha(4,0.5))
ecoforecastR::ciEnvelope(time,CI.fD[2,],CI.fD[4,],col=4)
lines(time,CI.fD[2,])
```

```{r}
## comparison to EnKF
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
lines(CI.fSSSD[2,],col="blue",lwd=2)
lines(Dbar,lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data","P(dist)"),
       lwd=5,col=c("orange","#b2df8a","blue","red","black"))
```

## Todo
-   Visualize bimodal forecast and analysis distributions through time for simple example
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA
